Verification of Stencil Computations through Spatial Specifications

### 2016-07-05 paper.tex
proof of IRREFL
progress on explaining the specification algebra used in the inference

\begin{definition}The product of two models of dimensionality
$n$ is given by the tensor $\otimes_n$ defined:
%%
\begin{alignat*}{2}
& \effdims{M}{n} =
\bigcup_{u \in M} \{x \mid x \in \{1,...,n\}, u_x \neq \infty\} \\
& M \otimes_n N =
  \Bigg\{x \; \Bigg| \;
    \parbox{5.6cm}{$i \in \{1, \ldots, 2^n\},$
                   $(u, v) \in M \times N,$ \\
                   $x = (u \bowtie v)_i,$ \\
                   $j \in (\effdims{M}{n} \cup \effdims{N}{n}),$ \\
                   $x_j \neq \infty$
                  } \Bigg\}
\end{alignat*}
The $\bowtie$ operation, which we call \emph{pairwise permutation}
(described below), constructs a $2^n \times n$ matrix ($2^n$-vector of
$n$-vectors). Using this, the $\otimes$ operation constructs the set of every
pairwise permutation for every pair $(u, v)$ of $n$-vectors in the Cartesian
product of $M \times N$.

@@

Given the set $\textsf{regions}(M)$ of spans covering the indexing
space, the next step converts these into abstract syntax trees
for the specification language.

For each span $S = [\vect{l} \; \vect{u}]$,
for each dimension, each pair of entries in the lower- and upper-bound
vectors is mapped to a concrete \textit{spec}, as defined by the
$\mathsf{conv}$ function:
%%
\begin{align*}
& \mathsf{conv}_i\; [\vect{l} \; \vect{u}] = \\
& \begin{cases}
%% FWD
 \stenFwd{\vect{u}_i}{i}{}  & \vect{l}_i = 0 \; \wedge \; \vect{u}_i > 0 \\[0.7em]
 \stenFwd{\vect{u}_i}{i}{, \irreflS}  & \vect{l}_i = 1 \; \wedge \; \vect{u}_i > 0 \\[0.7em]
%% BWD
 \stenBwd{|\vect{l}_i|}{i}{}  & \vect{l}_i < 0 \; \wedge \; \vect{u}_i = 0 \\[0.7em]
\stenBwd{|\vect{l}_i|}{i}{, \irreflS} & \vect{l}_i < 0 \; \wedge \; \vect{u}_i = -1 \\[0.7em]
%% CEN
 \stenCen{\vect{u}_i}{i}{}  &\vect{l}_i < 0 \; \wedge \; \vect{u}_i > 0
 \; \\
 & \qquad\quad \wedge \; |\vect{l}_i| = \vect{u}_i \\[0.7em]
 \stenBwd{|\vect{l}_i|}{i}{} & \vect{l}_i < 0 \; \wedge \; \vect{u}_i >
 0 \\
\quad \texttt{+} \; \stenFwd{\vect{u}_i}{i}{}  & \qquad\quad \wedge \; |\vect{l}_i| \neq \vect{u}_i \\[0.7em]
% ABSOLUTE REP
 \epsilon & \vect{l}_i = \infty \; \vee \; \vect{u}_i
 = \infty \\[0.7em]
% BOUND
 \texttt{atMost}, \stenFwd{\vect{u}_i}{i}{} & \vect{l}_i > 1 \\[0.7em]
 \texttt{atMost}, \stenBwd{|\vect{l}_i|}{i}{} \quad & \vect{l}_i < (-1) \\[0.7em]
  \end{cases}
\end{align*}
%%
The first two cases map to \texttt{forward} stencils since the lower
bound is $0$ or $1$, with the depth determined by the upper bound of
the span. If the lower bound is $1$, then the
$0$-offset point is missing so the specification is given an
\texttt{irreflexive} modifier. The second two cases are similar for
\texttt{backward} but now the depth of the region is determined by the
absolute of the lower bound. In the fifth-case, when the lower bound
is negative and the upper bound is positive and they have the same
magnitude, then a \texttt{centered} region is constructed with
the depth as the common magnitude. If the magnitudes are not the same
(sixth case) then separated \texttt{forward} and \texttt{backward}
regions are summed, with their corresponding depths.

In the case where either bound is $\infty$, representing a non
neighbourhood index, then a simple \texttt{irreflexive} specification
is given with no regions, indicating that there is not even reflexive
0-offset access in this dimension.

@@

\begin{lemma}
$\mathsf{conv}_i$ is total given the precondition that the lower
bound of the input span is less than or equal to its upper bound.
\end{lemma}

>>>>

\begin{definition}The product of two models of dimensionality
$n$ is given by the tensor $\otimes_n$ defined:
%%
\begin{alignat*}{2}
& \effdims{M}{n} =
\bigcup_{u \in M} \{x \mid x \in \{1,\ldots,n\}, u_x \neq \infty\} \\
& M \otimes_n N =
  \Bigg\{x \; \Bigg| \;
    \parbox{5.6cm}{$i \in \{1, \ldots, 2^n\},$
                   $(u, v) \in M \times N,$ \\
                   $x = (u \bowtie v)_i,$ \\
                   $j \in (\effdims{M}{n} \cup \effdims{N}{n}),$ \\
                   $x_j \neq \infty$
                  } \Bigg\}
\end{alignat*}
The $\bowtie$ operation, which we call \emph{pairwise permutation}
(described below), constructs a $2^n \times n$ matrix ($2^n$-vector of
$n$-vectors). Using this, the $\otimes$ operation constructs the set of every
pairwise permutation for every pair $(u, v)$ of $n$-vectors in the Cartesian
product of $M \times N$.

@@

Given the set $\textsf{regions}(M)$ of spans covering the indexing
space, the next step converts these into abstract syntax trees
for the specification language. This conversion is defined in terms
of an algebra on \textit{specTop} syntax which mirrors the shape
of the syntax with additive and multiplicative binary operators
$\oplus$ and $\odot$: 
%%
\begin{align*}
\oplus & : \textit{specTop} \rightarrow \textit{specTop} \rightarrow
  \textit{specTop} \\
\odot & : \textit{specTop} \rightarrow \textit{specTop} \rightarrow
        \textit{specTop} \\
\mathsf{toSpec} & : (\mathbb{Z} \cup \{\infty\}) \rightarrow (\mathbb{Z} \cup
  \{\infty\})  \rightarrow \textit{specTop}
\end{align*}
%%
The operation $\mathsf{toSpec}$ takes a pair of lower and upper bounds
values and produces a specification which contains a one-dimensional
region. The

$\oplus$ and $\odot$ respectively (defined in Definition~\ref{}
and Definition~\ref{} below).

The proceeds by converting each span
into a specification by converting ea
 is mapped to a specification,
and then summing specifications. This relies on product and sum
operations lifted to specification syntax (which is more than just
building regions out of \term{*} and \term{+}).

\begin{equation*}
\sum\limits_{\vect{x} \in \textsf{regions}(M)}
\prod\limits_{i \in \{1, \ldots, n\}} \, \mathsf{toSpec}(\vect{x}^L_i, \vect{x}^U_i)
\end{equation*}

For each span, and each dimension within it, each pair of lower and
upper bounds is mapped into a specification. That is,
for all $\vect{x}$ of size $n$, and for all $i \in \{1,\ldots,n\}$
then $\mathsf{toSpec} (\vect{x}^L_i, \vect{x}^U_i)$ maps each
pair of lower and upper bounds in each dimension to a piece of syntax 
of \textit{specTop} syntax, defined:
%%
%\newcommand{\llower}{\vect{x}^L_i}
%\newcommand{\lupper}{\vect{x}^U_i}
\newcommand{\llower}{l}
\newcommand{\lupper}{u}
\begin{align*}
&\mathsf{toSpec} \; (\llower, \lupper) = \\
&\begin{cases}
%% FWD
 \stenFwdS{\lupper}{i}  & \llower = 0 \, \wedge \, \lupper > 0 \\[0.3em]
 \stenFwdSR{\lupper}{i}{\irreflS}  & \llower = 1 \, \wedge \, \lupper > 0 \\[0.3em]
%% BWD
 \stenBwdS{|\llower|}{i}  & \llower < 0 \, \wedge \, \lupper = 0 \\[0.3em]
\stenBwdSR{|\llower|}{i}{\irreflS} & \llower < 0 \, \wedge \, \lupper = -1 \\[0.3em]
%% CEN
 \stenCenS{\lupper}{i}  & \llower < 0 \, \wedge \, \lupper > 0
 \, \wedge \, |\llower| = \lupper \\[0.3em]
 \stenBwdS{|\llower|}{i} & \llower < 0 \, \wedge \, \lupper > 0  \, \wedge \, |\llower| \neq \lupper \\
\quad \texttt{+} \, \stenFwdS{\lupper}{i}  & \qquad\quad \\[0.3em]
% ABSOLUTE REP
 \epsilon & \llower = \infty \, \vee \, \lupper
 = \infty \\[0.3em]
% BOUND
 \texttt{atMost}, \stenFwdS{\lupper}{i} & \llower > 1 \\[0.3em]
 \texttt{atMost}, \stenBwdS{|\llower|}{i} \quad & \llower < (-1) \\[0.3em]
  \end{cases}
\end{align*}
%%
The first two cases map to \texttt{forward} stencils since the lower
bound is $0$ or $1$, with the depth determined by the upper bound of
the span. If the lower bound is $1$, then the
$0$-offset point is missing so the region is given an
\texttt{irreflexive} attribute. The second two cases are similar for
\texttt{backward} but now the depth of the region is determined by the
absolute of the lower bound. In the fifth-case, where the lower bound
is negative and the upper bound is positive and they have the same
magnitude, then a \texttt{centered} region is constructed with
the depth as the common magnitude. If the magnitudes are not the same
(sixth case) then separated \texttt{forward} and \texttt{backward}
regions are summed, with their corresponding depths.

In the case where either bound is $\infty$, representing a non
neighbourhood index, then an empty specification is given
 with no regions, indicating that there is not even reflexive
0-offset access in this dimension.

@@

\begin{lemma}
$\mathsf{toSpec}_i$ is total (\ie{}, defined for all spans) given the precondition that the lower
bound of the input span is less than or equal to its upper bound.
\end{lemma}

\begin{equation*}
{\Large{\bullet}}_{i \in \{1, \ldots, n\}} \, \mathsf{toSpec}(\vect{x}^L_i, \vect{x}^U_i)
\end{equation*}
